The Rockerverse: Images, Packages, and Applications for Containerization with R

### 2020-01-26 manuscript.Rmd
added some 'dockr' remarks

@@

**\pkg{dockerfiler}** ([https://github.com/ColinFay/dockerfiler/](https://github.com/ColinFay/dockerfiler/)) is an R package designed for building `Dockerfile`s straight from R.
A scripted creation of a `Dockerfile` enables iteration and automation, for example for packaging applications for deployment (see \nameref{deployment}).
Being scriptable from R developers can leverage the tools available in R to parse a `DESCRIPTION` file, to get system requirements, to list dependencies, versions, etc.
**\pkg{containerit}** ([https://github.com/o2r-project/containerit/](https://github.com/o2r-project/containerit/)) attempts to take this one step further and includes these tools to automatically create a `Dockerfile` that can execute a given workflow \citep{nust_containerit_2019}.
\pkg{containerit} accepts R `sessionInfo` objects as input and provides helper functions to derive these from workflows, e.g., an R script or R Markdown document, by analysing the session state at the end of the workflow.
It relies on the \pkg{sysreqs} ([https://github.com/r-hub/sysreqs/](https://github.com/r-hub/sysreqs/)) package and it's mapping of package system dependencies to platform specific installation package names\footnote{See \href{https://sysreqs.r-hub.io/}{https://sysreqs.r-hub.io/}.}.
\pkg{containerit} uses \CRANpkg{stevedore} to streamline the user interaction and improve the created `Dockerfile`s, e.g., by running a container for the desired base image to extract the already available R packages.
**\pkg{dockr}** ([https://github.com/smaakage85/dockr](https://github.com/smaakage85/dockr)) is a very similar package attempting to mirror a given R session, including local non-CRAN packages\citep{kjeldgaard_dockr_2019}.
Users can manually add statements for non-R dependencies to the `Dockerfile`.
**\CRANpkg{liftr}** aims to solve the problem of persistent reproducible reporting in statistical computing based on the R Markdown format \citep{xie2018} for dynamic documents \citep[\href{https://nanx.me/liftr/}{https://nanx.me/liftr/}, ][]{liftr2019}.
The irreproducibility of authoring environments can become an issue for collaborative documents and large-scale platforms for processing documents.
\CRANpkg{liftr} makes the document the main and sole workflow control file and the only file that needs to be shared between collaborators for consistent environments, e.g. demonstrated in the DockFlow project (https://dockflow.org).
It introduces new fields to the R Markdown document header, allowing users to manually declare the dependencies, including versions, for rendering the document.
The package then generates a `Dockerfile` from this metadata and provides a utitility function to render the document inside a Docker container, i.e., `render_docker("foo.Rmd")`.
An RStudio addin even allows compilation of documents with a single push of a button.

>>>>

  - name: Lars Kjeldgaard
    # https://github.com/smaakage85
    affiliation: Danish Tax Authorities
      - Oestbanegade 123
      - 2100, Koebenhavn Oe
    email: lars_kjeldgaard@hotmail.com

@@

**\pkg{dockerfiler}** ([https://github.com/ColinFay/dockerfiler/](https://github.com/ColinFay/dockerfiler/)) is an R package designed for building `Dockerfile`s straight from R.
A scripted creation of a `Dockerfile` enables iteration and automation, for example for packaging applications for deployment (see \nameref{deployment}).
Being scriptable from R developers can leverage the tools available in R to parse a `DESCRIPTION` file, to get system requirements, to list dependencies, versions, etc.
**\pkg{containerit}** ([https://github.com/o2r-project/containerit/](https://github.com/o2r-project/containerit/)) attempts to take this one step further and includes these tools to automatically create a `Dockerfile` that can execute a given workflow \citep{nust_containerit_2019}.
\pkg{containerit} accepts R `sessionInfo` objects as input and provides helper functions to derive these from workflows, e.g., an R script or R Markdown document, by analysing the session state at the end of the workflow.
It relies on the \pkg{sysreqs} ([https://github.com/r-hub/sysreqs/](https://github.com/r-hub/sysreqs/)) package and it's mapping of package system dependencies to platform specific installation package names\footnote{See \href{https://sysreqs.r-hub.io/}{https://sysreqs.r-hub.io/}.}.
\pkg{containerit} uses \CRANpkg{stevedore} to streamline the user interaction and improve the created `Dockerfile`s, e.g., by running a container for the desired base image to extract the already available R packages.
**\pkg{dockr}** ([https://github.com/smaakage85/dockr](https://github.com/smaakage85/dockr)) is a similar package. It is often desirable to organize R code in the R package structure. \pkg{dockr} facilitates the creation of a Docker container image for your R
package, in which the package itself is available. The container image mirrors your current R session and includes all of the R dependencies needed to run your R package - also local non-CRAN packages\citep{kjeldgaard_dockr_2019}. Users can manually add statements for non-R dependencies to the `Dockerfile`.
**\CRANpkg{liftr}** aims to solve the problem of persistent reproducible reporting in statistical computing based on the R Markdown format \citep{xie2018} for dynamic documents \citep[\href{https://nanx.me/liftr/}{https://nanx.me/liftr/}, ][]{liftr2019}.
The irreproducibility of authoring environments can become an issue for collaborative documents and large-scale platforms for processing documents.
\CRANpkg{liftr} makes the document the main and sole workflow control file and the only file that needs to be shared between collaborators for consistent environments, e.g. demonstrated in the DockFlow project (https://dockflow.org).
It introduces new fields to the R Markdown document header, allowing users to manually declare the dependencies, including versions, for rendering the document.
The package then generates a `Dockerfile` from this metadata and provides a utitility function to render the document inside a Docker container, i.e., `render_docker("foo.Rmd")`.
An RStudio addin even allows compilation of documents with a single push of a button.
